Update\_Writeup

### 2019-02-27 R Files/Updates_Writeup.Rmd
Finished EDA for contrast and effective dose, knitted to pdf

We are also interested in the relationship of the effecive dose and contrast administered with both the age and BMI of the patient. Patients were divided into 10-year age groups and 5-unit BMI groups. In both cases, cut-off values were determined based on the number of patients that fell into the extreme groups. For example, there was only one patient in the data under the age of 10, so they were added into the "<20" group. The number of patients that belong to each age group and BMI group is given in Table 2 and 3 respectively. Figure 3 shows the average radiation administered for each of these age groups.

```{r}
summary(PE$Age10)
rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Figure 3: Mean Effective Dose by Age", ylim = c(3,7))
points(rad_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)
```

## Contrast Given

>>>>

We are also interested in the relationship of the effecive dose and contrast administered with both the age and BMI of the patient. Patients were divided into 10-year age groups and 5-unit BMI groups. In both cases, cut-off values were determined based on the number of patients that fell into the extreme groups. For example, there was only one patient in the data under the age of 10, so they were added into the "<20" group. The number of patients that belong to each age group and BMI group is given in Table 2 and 3 respectively.

\begin{table}[H] \centering 
  \caption{Breakdown Patients by Age Group} 
\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 <20 & 20-30 & 30-40 & 40-50& 50-60 & 60-70 & 70-80 & 80-90 & >90 \\ 
 \hline \\[-1.8ex] 
55 & 133 & 222 & 294 & 514 & 562 & 397 & 179 & 26 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

\begin{table}[H] \centering 
  \caption{Breakdown Patients by BMI Group} 
\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 <15 & 15-20 & 20-25 & 25-30 & 30-35 & 35-40 & 40-45 & 45-50 & 50-55 & 55-60 & >60 \\ 
 \hline \\[-1.8ex] 
14 & 158 & 486 & 595 & 467 & 276 & 171 & 106 & 56 & 21 & 32 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

Figure 3 shows the average radiation and contrast administered for each of the age groups, and Figure 4 shows the average radiation and contrast administered for the BMI groups.

```{r, fig.height = 5, fig.width = 10}
par(mfrow = c(1,2))
rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Figure 3a: Mean Effective Dose by Age", ylim = c(3,7), cex.main = 1)
points(rad_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))

contrast_by_age <- tapply(PE$`Contrast Given`, PE$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Figure 3b: Mean Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(75,95))
points(contrast_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
axis(2, at = seq(75,95,5))
par(mfrow = c(1,1))
```

```{r, fig.height = 5, fig.width = 10}
par(mfrow = c(1,2))
rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
     main = "Figure 4a: Mean Effective Dose by BMI", cex.main = 1)
points(rad_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))

contrast_by_bmi <- tapply(PE$`Contrast Given`, PE$BMI10, "mean")
plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
     main = "Figure 4b: Mean Contrast Administered by BMI", cex.main = 1)
points(contrast_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
par(mfrow = c(1,1))
```

Interestingly, we see nearly identical patterns for both the radiation and contrast administered when divided into either age groups or BMI groups. For age, we see a nearly quadratic relationship, with more radiation and contrast being administered for groups between 30 and 60 years of age, and lower amounts administered to both younger and older groups.

On the other hand, we have seem to have a purely increasing relationship for radiation and contrast with BMI. As a patient's BMI increases, we can see that the average effective dosage and amount of contrast given both increase.

From these results, it appears that the amount of radiation and contrast administered to a patient are correlated. Indeed, we find a Pearson's correlation of 0.47, which is significant (t(2380) = 26.05, p < .001).

### 2019-02-27 R Files/Updates_Writeup.tex
Finished EDA for contrast and effective dose, knitted to pdf

\subsection{Effective Dose}\label{effective-dose}

@@

Using the fixed data for effective dosage of radiation administered to
patients, we retrieve the plots shown in Figure 1. Here we can see the
right-skew in the histogram suggesting several patients recieved much
higher amounts of radiation than average. The mean effective dose given
to a patient was 5.631 millisieverts, with a range of 0.1 to 48.8
millisieverts. Out of all 2382 patients, only 54 recieved a radiation
dose above 20 millisieverts.

@@

We are also interested in the relationship between the effecive dose
administered and both the age and BMI of the patient. For the case of
age, patients were grouped into 10-year age groups. The number of
patients that belong to group is given below. Figure 2 shows the average
radiation administered for each of these age groups.

\begin{verbatim}
##   <20 20-30 30-40 40-50 50-60 60-70 70-80 80-90   >90 
##    55   133   222   294   514   562   397   179    26
\end{verbatim}

@@

>>>>

\subsection{Effective Dose and
Contrast}\label{effective-dose-and-contrast}

@@

Numerical summaries of both effective dose and contrast are given in
Table 1. Using the fixed data for effective dosage of radiation
administered to patients, we retrieve the plots shown in Figure 1. Here
we can see the right-skew in the histogram suggesting several patients
recieved much higher amounts of radiation than average.

@@

We also have constructed an updated histogram and boxplot for the amount
of contrast given to patients, which are shown in Figure 2. Here we see
a centralized distribution with some extreme values on both sides.

@@

We are also interested in the relationship of the effecive dose and
contrast administered with both the age and BMI of the patient. Patients
were divided into 10-year age groups and 5-unit BMI groups. In both
cases, cut-off values were determined based on the number of patients
that fell into the extreme groups. For example, there was only one
patient in the data under the age of 10, so they were added into the
``\textless{}20'' group. The number of patients that belong to each age
group and BMI group is given in Table 2 and 3 respectively.

\begin{table}[H] \centering 
  \caption{Breakdown Patients by Age Group} 
\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 <20 & 20-30 & 30-40 & 40-50& 50-60 & 60-70 & 70-80 & 80-90 & >90 \\ 
 \hline \\[-1.8ex] 
55 & 133 & 222 & 294 & 514 & 562 & 397 & 179 & 26 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

\begin{table}[H] \centering 
  \caption{Breakdown Patients by BMI Group} 
\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 <15 & 15-20 & 20-25 & 25-30 & 30-35 & 35-40 & 40-45 & 45-50 & 50-55 & 55-60 & >60 \\ 
 \hline \\[-1.8ex] 
14 & 158 & 486 & 595 & 467 & 276 & 171 & 106 & 56 & 21 & 32 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

Figure 3 shows the average radiation and contrast administered for each
of the age groups, and Figure 4 shows the average radiation and contrast
administered for the BMI groups.

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-4-1.pdf}

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-5-1.pdf}
